---
title: "PS166_Section11"
author: "Johanna Reyes O."
date: "2023-11-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) #this one includes both dplyr and ggplot2!!!
```

## Basics recap, using the "anes_latino" data 

Today, we will be using the following variables:
* The prevote_voted variable notes whether a responded voted in the last General Election. 
* The trustgov_trustgrev variable stands for trust in government.
* The gender_respondent_x is a binary measure of gender.
* The dem_hisptyp_1st variable is the first mention of country of origin.
* The dem_racecps_white refers to whether respondents consider their race to be white.

### 1. How do we count total number of observations (i.e. the total number of respondents) within a dataframe?

```{r}
#read data- I am using the RDS fie below
anes <- readRDS("anes.rds")

#total observations (we know two ways!)
```

### 2. How do we know for how many respondents we have missing information about one variable? (check first prevote_voted, then trustgov_trustgrev) 

```{r}
#NAs for voting

#NAs for trust

```
### 3. How do we know for how many respondents we have information about one variable? (check first prevote_voted, then trustgov_trustgrev) 

```{r}
#info for voting


#info for trust 

# Alternative: count number of total observations and subtract missing ones 
```
### 4. How do we know for how many respondents we have information for two variables? (check for dem_hisptype_1st AND prevote_voted)

```{r}
sum(!is.na(anes$var1) & !is.na(anes$var2)) #substitute var1 and var2 with correct variable names

```
### 4. Compute a bar plot with the MEDIAN value of trust in government (in the y-axis) by gender (in the x-axis). Order the bars from the group with the highest to the lowest trust in government.

* Remember: when computing statistics, for example, a mean or median, we need to include "na.rm=TRUE" to exclude missing cases. Otherwise, we will get NA as the value of the mean.

Let's recall how to use ggplot:

```{r}
# A basic barplot of trust only takes the x variable, but we want the MEDIAN trust BY Group (Gender)

# Basic plot 

ggplot(data = anes, aes(x = trustgov_trustgrev)) + #first later specifying data and variables
  geom_bar() #second layer specifying plot type

# What is the first thing we need to clean based on the plot?
  
```
Let's now plot trust levels by group
```{r}
# First, clean and group the data

anes %>% # this tell's R to use the data 
  filter(!is.na(trustgov_trustgrev)) %>% # keep only what is NOT missing in trustgov_trustgrev variable
  group_by(gender_respondent_x) %>% # we want average TRUST by GENDER, so group by gender first
  summarize(med_trust = median(trustgov_trustgrev, na.rm = TRUE)) %>%  # Summarize by the median. We create a new variable so we can use that as the y axis
ggplot(aes(y = med_trust , x=  gender_respondent_x)) + # specify variables
    geom_col() #specify type of plot

```
Now, repeat but plot the mean trust levels by Hispanic origin type. Add code to the LEFT of each comment. Remember that we use piping %>% to manipulate the data and then we add ggplot and its layers  

```{r}
 # this tell's R to use the data 
 # keep only what is NOT missing in trustgov_trustgrev variable
 # we want average TRUST by GENDER, so group by gender first
 # Summarize by the median
 # specify variables
 #specify type of plot
```


```{r}

#Now, together reorder by decreasing mean trust levels, add fill color, and labels for axes and each bar using "+ geom_text(aes(label = round(mean_trust, 2)))"

  
```

### 5. 

* What is the average value of trust in government for Puerto Ricans?
* What is the average value of trust in government for Chicanx? 

```{r}

## 6. Let's now make a density plot for trust levels by gender

# basic density plot 

anes %>%
  filter(!is.na(trustgov_trustgrev)) %>% 
  ggplot(aes(x = trustgov_trustgrev))+
  geom_density()

# now separate by gender- note we do not need to group by because we are not calculating any statistics- we can just color differently

anes %>%
  filter(!is.na(trustgov_trustgrev)) %>%
  ggplot(aes(x = trustgov_trustgrev, fill = factor(gender_respondent_x))) + #we use fill to add a fill color by gender
  geom_density(alpha = 0.5) + # alpha specifies a transparency level
  labs(fill = "Gender( 1= Male, 2=Female)", x= "Mean Trust Levels") #labs allows to specify multiple labels

```

### 6. Let's compute a scatter plot displaying the correlation between levels of trust (in the x-axis) whether the respondent voted in the General Election. Let's include a linear regression summarizing the correlation.
```{r}
anes %>%
   filter(!is.na(trustgov_trustgrev)) %>%
  ggplot(aes(x = trustgov_trustgrev, y = prevote_voted))+
  geom_point()+
 geom_smooth(method=lm) 

#How can we look at the actual regression estimation?

#summary(lm(y~x, data = data))
```
### 7. Now, let's repeat but identifying with a different color Hispanics by gender.

```{r}
anes %>%
   filter(!is.na(trustgov_trustgrev)) %>%
  ggplot(aes(x = trustgov_trustgrev, y = prevote_voted, color = factor(gender_respondent_x))) + #note how we use 'color' and not 'fill' because it is a line (but the idea is the same as above!)
  geom_point()+
 geom_smooth(method=lm)

#Add labels to the plot

#Discuss with the person next to you how you would interpret the plot
 

```

Repeat but now color by their first mention of race is white 'dem_racecps_white'

```{r}
#first use table() to see who the variable values are distributed

#plot a scatterplot, add a regression line, and appropriate labels 
```


