---
title: "PS 166 Section 8"
output: pdf_document
date: "2023-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

IRkernel::installspec()

```

# Section 8

Today, you will be working through this RMD with a partner and then each person will upload their file to the group responses folder on Bcourses. 

### Review the following key concepts with your partner

* What is a data frame?
** What do columns and rows represent?
* What is a variable?
* What is an observation?
* What is an object?
* What is a command?


Recall from last week that in order to understand what the variable names and values mean, it is vital to understand the related codebook. A codebook provides a description of each variable, including its name, label, data type, and values. It also includes information about how the variable was measured, how missing values are coded, and any special codes or categories used in the variable. 

* Why might researchers want to create a thorough and clear codebook?

### Let's dive into the ANES codebook

This dataframe contains many variables! Of these, we might not be interested in all. For today, we will focus on variables related to voting and demographic characteristics; it is a Latinx politics course after all...

Open the codebook. This should already be inside your Jupyter Hub directory. 

Now, scroll through the codebook and extract the following important details:

*  What is the population?
*  What is a sample?
*  Why are the data weighted?

### Understanding variables of interest

Now that we know the connection between the codebook and the data, let's gather some information about Latinx politics. Discuss the following with a partner.

* What kind of demographic data did the survey collect?
* What is the percentage of people who are Spanish, Hispanic or Latino?
* What countries of origin do respondents come from?
** Where do these categories come from?
* What kind of political preferences and participation did the survey collect?

Time to play around with two variables of interest. Let's say we are interested in evaluating how many Latinx/os/as voted. Which variables should we look at?


What kind of variables are these? (Categorical, ordinal, continuous, etc.)

## Checkpoint 1

### Working with Missing Data

There are at least two reasons why we may want to know whether our data has missing values (i.e. NAs). First, it’s important to have a sense of how large our sample is. Second, some functions in R, like the computation of an average, are sensitive to the inclusion of missing values (NAs). In this case, we may want to be aware of whether we have missing values to deal with them appropriately. We can either filter out missing values, or indicate in the function that missing values should be ignored. 

For example, in the case of a computation of a mean of variable x you would use the following command:  `mean(x, na.rm = TRUE)`, where na.rm = TRUE, tells R to ignore the NAs and to compute the average value with the non-missing observations. Below we will practice how to filter out missing values.

```{r}
#df$variable of interest

#Refer to the dem_hisp and postvote_rvote variables using the $ symbol. Make sure to load the data first using the 
#'read.csv("name of file.csv")' command.

#Using the command we used last week, count the number of NAs #we have (1) in the entire dataset, 
#(2) in the "Hispanic type - 1st mention" variable


#Now, count for how many observations we have data in each variable. Hint: use #the 'sum' command and
#is.na, plus the negation sign '!'


#Now, assess the distribution of the variables. That is, what is the number of observations in each 
#of the different categories of the variables. Use the command 'table' and refer to one specific 
#variable at a time. 

```

* What is the modal category (that is, the category with the largest number of observations, or the most frequent category) for the two variables?
 
```{r}
#Check the range of the two variables using the 'range' command


#How do we interpret this range? Let's take a look at the codebook again.

```

Let's learn how to filter unwanted data.

```{r}

#First, let's load the package "dyplr" by running the following code

library(dplyr)

#Now, let's filter out any NA values from the anes dataframe. The %>% operator #in R comes from the dplyr package 
#and allows you to pipe the output of one function into the input of another function.

#First, let's recall the distribution of the postvote_rvote variable with the command 'table'


#Now, let's filter out missing values
anes %>%
  filter(!is.na(postvote_rvote))

#Look at the distribution of the variable again... any changes?


```

Here, we are telling R to keep only those values in the postvote_rvote column that ARE NOT missing or NAs. Note that we do not need to re-specify the data as we did before (anes$postvote_rvote) because we have already specified the dataframe through the piping operator.

If you do it this way, any row that has a missing value for the postvote_rvote variable will be deleted, including values in other columns. Therefore, if we want to retain those values, we might considering assigning the filtered dataframe to a new object, without modifying the original dataframe. How would you do this?

```{r}

#Count the total observations in each variable 


#Retry the above but assign the filtered dataframe to a new object 

```

```{r}
#Example: Let's try again with a different variable 'strata_ftf'. Display it's distribution first, then filter 
#out missing values

table(anes$strata_ftf)
sum(is.na(anes$strata_ftf))

anes <- anes %>%
  filter(!is.na(strata_ftf))

table(anes$strata_ftf)
```

## Checkpoint 2

Now, let's learn how to compute statistics. 


Is there a relationship between gender and voting? One way to assess whether there is a relationship between gender and voting is to compare the average turnout of men to the average turnout of women. If these averages are very different to each other, then we would conclude that there is a relationship between gender and voting.

Find the mean of postvote_rvote for women-identifying respondents only. First, write down how a mean is calculated. 

First, find the mean of postvote_rvote for the entire sample. Use:

mean(data$var1, na.rm = TRUE)

```{r}

```

Then, use the following syntax to find the mean for women-identifying individuals only:

mean(data$var1[data$var2 == ?], na.rm = TRUE)


```{r}

```

Find the mean of postvote_rvote for men-identifying respondents only.

```{r}

```

What does na.rm do in this case? Is it applicable here? Why/why not?


Let's use the piping operator to find the mean of the entire voting variable.

```{r}

anes %>%
  summarize(mean(postvote_rvote), na.rm = TRUE)

#Repeat for the median, standard deviation (sd), minimum and maximum value, and percentiles. Before computing the 
#minimum and maximum, recall that we have already learned one way to do so...


#Note: for percentiles, you need to specify which (e.g. 0.25, 0.75, etc.)



#You can also use piping it to count missing values! Try below.


```

## Checkpoint 3

Now, let's repeat computing the same statistics, but by grouping the data by Hispanic origin. This is where the magic of piping really comes through...

```{r}
#Find the mean of the voting variable by Hispanic origin

anes %>%
  group_by(dem_hisptyp_1st) %>%
  summarize(mean(postvote_rvote))
```

Repeat for the median, standard deviation (sd), minimum and maximum value, and 75th percentile.

Mean: The "average" number; found by adding all data points and dividing by the number of data points.

Median: The middle number; found by ordering all data points and picking out the one in the middle (or if there are two middle numbers, taking the mean of those two numbers).

The 75th percentile is the value at which 25% of the answers lie above that value and 75% of the answers lie below that value.

```{r}


```
Now, let's combine (1) removing NAs, (2) summarizing by Hispanic origin group, and (3) computing the mean of the highest level of education of the respondents using the piping method. Assign the result to a new object.

```{r}


```

