---
title: "PS 231B Section 5- Exercise"
author: "Your Name"
date: "February 14, 2024 <3"
output:
  pdf_document
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#set chunk options
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# load libraries using pacman
if (!require("pacman")) install.packages("pacman")
pacman::p_load(png, grid, ggplot2, tidyverse, MASS, knitr, Matching, stargazer, optmatch, RItools)
```

# Review of concepts

In pairs, define the following and discuss (1) when/how is it useful? and (2) an example of a possible application:

- Common support 
- Propensity scores
    + In relation to curse of dimensionality?
    + In relation to inverse probability weighting?

- Issues faced when using linear regression to  estimate treatment effects

# Balance tests: Recap

Let's first refresh our memories by checking balance after we matched with the Euclidean distance last week...

Recall the following: 

Arrington et al. (2021) evaluate whether the processes states use to select judges for high courts influence gender diversity. They use a global sample sine 1970 and a matching design in two stages (exact matching and propensity score matching) to compare treated and control units with the same pretreatment institutions. Treatment is a change in the number of appointers or nominators.

[Here is the paper.](https://www-cambridge-org.libproxy.berkeley.edu/core/journals/american-political-science-review/article/constitutional-reform-and-the-gender-diversification-of-peak-courts/D9B56B946416D4582B34D769B3F937F5#supplementary-materials) We will be using only  `women_highcourts.csv`. 

(If curious, you can find the authors' code and original data files [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/RHPSZD)). 

```{r}

# Load the data

```

Let's match by Euclidean distance first on the same variables as last time, and check balance for them...Hint: simply copy-paste your code from last week. 

We are interested in the effect of 
`"treated1"`: or treatment as an increase in the number of nominators/selectors on `"has_there_ever_been_wom"`: which indicates time until the year of appointment of the first woman to a state’s highest court (ie. 0 in the years prior to appointment and 1 once the first woman is appointed). In the first stage, the authors only match on:

>- `year`
>- `nominators_to_match1` 
>- `selectors_to_match1`

# Matching as before

```{r}
# missing values in outcome, and Match does not like that...so subset data to only include needed columns and omit NAs in the outcome variable


# define X

# define Y

# define treatment 

# perform matching 

# let's now check balance before and after matching

```

# Now, with propensity score

What are the necessary steps? 



For the propensity score weighting, the authors actually take the matched dataset from step 1 and then match again using a different set of variables.

Comment the code below:
```{r}
#First, we need to extract the matched data from the above

#summary.Match(euclidean_matches)

matched <- euclidean_matches$mdata

index.treated <- euclidean_matches$index.treated
index.control <- euclidean_matches$index.control

# # Create match_id based on order of indexes  
match_id <- c(1:length(index.treated), 1:length(index.control))

# Create match id for full data- not just matched observations

full_match_id <- rep(0, nrow(X))

# Assign match ids for those that were matched
full_match_id[order(c(index.treated, index.control))] <- match_id

# Bind back to FULL original data  
new_data <- as.data.frame(cbind(Y, X, Tr,  
                  MatchID = full_match_id))

# merge with all other covariates

## to ensure merging works- we need to make sure the original df has the same dimensions and column names

names(new_data)[c(1, 5)] <- c("has_there_ever_been_wom", "treated1")

new_data_cols <- names(new_data)[1:5]

# Merge with all covariates
highcourts2 <- merge(new_data, highcourts1, by = new_data_cols)

# delete duplicates
highcourts2 <- highcourts2 %>% 
  distinct(record_id, .keep_all = TRUE)

#write.csv(highcourts2, "highcourts2.csv", row.names = F) # 

```

# Reproducing Propensity Score Matching from Arrington et al. 2021

**Comment the authors' code below WITHOUT ACTUALLY RUNNING IT!** 

```{r, warning=FALSE, echo=FALSE}
# #install.packages("optmatch")
# #install.packages("RItools")
# # library(optmatch)
# # library(RItools)
# 
# #propensity score matching with optmatch
# psm<-glm(treated1~ percent_women_filled 
#          +since_UNIVSFFRG   
#          +v2jupack_filled
#          +v2x_partipdem_filled
#          +v2x_egaldem_filled
#          + v2xcl_rol_filled
#          +v2x_gender_filled
#          +v2csgender_filled
#          +v2mefemjrn_filled
#          +v2cldmovew_filled
#          +v2cldiscw_filled
#          +v2clslavef_filled
#          +v2clprptyw_filled
#          +v2clacjstw_filled
#          +v2pepwrgen_filled
#          +strata(subclass), family="binomial", data=cem.data2017_tr1)
# 
# ps.pm <- pairmatch(psm, data = cem.data2017_tr1)
# 
# summary(ps.pm)
# 
# 
# #pulling matches indicator from optmatch into the dataframe
# all.equal(names(ps.pm), row.names(cem.data2017_tr1)) #TRUE
# matched_pairs_tr1<-cbind(cem.data2017_tr1, matches=ps.pm)

```

# An aside: Why GML? 

A Generalized linear model (GLM) is a generalization of ordinary linear regression that allows for response variables that have an error distribution other than a normal distribution... hence, more "general"...

GLMs also have a non-linear 'link functions', which links the regression coefficients to the distribution and allows the linear model to generalize.

Useful resources:
>- [Datacamp Module](https://www.datacamp.com/tutorial/generalized-linear-models) and [Interactive Course](https://app.datacamp.com/learn/courses/generalized-linear-models-in-r)
>- [`gml` R documentation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/glm)
  >- Including foundational references
  
>- More in 231C!

# Back to matching...

Now, modify the authors' code so that we use our new_data df and the matchID as a covariate instead of their strata variable....

```{r}
#propensity score matching with optmatch


```
Now, let's practice plotting our propensity scores... fill in the "blanks":

```{r}

# get the propensity scores from the model
## note, we add the original df to be able to merge - this is code from lecture

#define treatment for plotting - to create a new df with pscores and the treatment

# Plot pscores, color by treatment group

# now, let's match like the authors do
ps.pm <- pairmatch(psm, data = highcourts2)
summary(ps.pm) # comment on the results

#pulling matches indicator from optmatch into the dataframe
all.equal(names(ps.pm), row.names(highcourts2)) #TRUE
matched_pairs_tr1<-cbind(highcourts2, matches=ps.pm)

# Refit glm on matched data using the same model as above - just so we can plot with the matched data- the pscores should not change

# Predict propensity scores

#Keeping version with all data for generating balance plot later
matched_pairs_tr1_full<-matched_pairs_tr1

# plot again- but only using the matches!!! 

# How effective were the new matches?
```

# Balance Plots

Now, let's check balance again to see if our matching strategy worked

```{r}
####Balance#### from the authors' code
# Comment as you run!


#make matched pairs numeric
matched_pairs_tr1_full$matched<-as.numeric(matched_pairs_tr1_full$matches)
matched_pairs_tr1_full$matched[is.na(matched_pairs_tr1_full$matched)] <- 0


#need a vector of the variables that we need
var_names<-   c("year",
                "nominators_to_match1",
                "selectors_to_match1", 
                "percent_women_filled",
                "since_UNIVSFFRG",
                "v2jupack_filled",
                "v2cldiscw_filled",
                "v2clprptyw_filled",
                "v2csgender_filled",
                "v2pepwrgen_filled",
                "v2x_gender_filled",
                "v2xcl_rol_filled",
                "v2clacjstw_filled",
                "v2cldmovew_filled",
                "v2clslavef_filled",
                "v2mefemjrn_filled",
                "v2x_egaldem_filled",
                "v2x_partipdem_filled")
matched_sdm<-NULL
all_sdm<-NULL
balance<-as.data.frame(cbind(var_names, matched_sdm, all_sdm))


for (i in 1:length(var_names)){
  balance$matched_sdm[i]<- (((mean(na.omit(matched_pairs_tr1_full[matched_pairs_tr1_full$treated1==1, var_names[i]]))) 
                             -
                               (mean(na.omit(matched_pairs_tr1_full[matched_pairs_tr1_full$treated1==0 & matched_pairs_tr1_full$matched!=0, var_names[i]]))))
                            /
                              (sqrt(var(na.omit(matched_pairs_tr1_full[matched_pairs_tr1_full$matched!=0, var_names[i]])))))
}

for (i in 1:length(var_names)){
  balance$all_sdm[i]<- (((mean(na.omit(highcourts[highcourts$treated1==1, var_names[i]]))) 
                         -
                           (mean(na.omit(highcourts[highcourts$treated1==0, var_names[i]]))))
                        /
                          (sqrt(var(na.omit(highcourts[, var_names[i]])))))
}


balance_variables<-c("Year", 
                     "Number of Nominators", 
                     "Number of Selectors",
                     "Percent Women in Parliament", 
                     "Years since Univ. Suffrage", 
                     "Court Packing Index", 
                     "Freedom of Discussion for Women", 
                     "Property Rights for Women", 
                     "Women's Civil Soc. Participation", 
                     "Power Distributed by Gender",
                     "Women Political Empowerment", 
                     "Equality Before the Law", 
                     "Access to Justice for Women", 
                     "Freedom Domestic Mvmt for Women", 
                     "Freedom from Forced Labor for Women", 
                     "Female Journalists", 
                     "Egalitarian Democracy Index",
                     "Participatory  Democracy Index" 
)



#PLOT #
#pdf("Balance_Plot_Tretment_as_Increase_jan2021.pdf", height=5, width=10 ) 
par(mfrow=c(1,1), mar=c(5, 18, 5, 5), family="serif")
y.axis <- c(length(balance$matched_sdm):1)
plot(balance$matched_sdm, y.axis,
     type = "p",
     axes = FALSE,
     xlab = "",
     ylab = "",
     pch = 20,
     cex = 1.2,
     xlim = c(-1,1),
     main = "Standardized Difference in Means",
     sub="Difference between Treated and Untreated Means"
)

segments(0, 0, x1=0, y1=19, lty = 2, lwd = 2, col = "grey")
axis(2, at = y.axis, label = balance_variables, las = 1, tick = T, mgp = c(2,.6,0), cex.axis = 1)
axis(1, mgp = c(0,.6,.75))
points(balance$all_sdm, y.axis, pch=1, cex=1)
legend(x= .2, y=18, c("Matched Data", "All Data"), pch=c(20, 1), cex=.8,  pt.cex=.8, bty = "n")
mtext("Balance, Treatment as Increase")
#dev.off()

# Take a look at the paper-- did we effectively reproduce this table?

```

Did propensity score give us better balance than Euclidean distance matching? Why should it in this example?

# IPW Estimator

Now, if you got this far,  manually calculate the difference in means with IPW...

This requires two steps:
1. 
2. 

```{r}

# Use only matched data

# Calculate weights

# Divide each unit’s weight by the sum of all weights in that group to normalize weights- why do we do this?

# Weighted means 

# Estimate difference in means                            

# Print                             


#Is this an estimate for the ATE or ATT????
```

# If you got this far...        

Repeat with all observations

```{r}
# Calculate weights

# Divide each unit’s weight by the sum of all weights in that group

# Weighted means  


# Estimate DIM

```

How could we deal with this issue? (Hint: check out Lec. 5 slide 20)...
